Modelling the City of Melbourne Pedestrian Data

### 2017-09-29 presentation/final.Rmd
update

```{r, echo = FALSE, fig.height=5, fig.retina = 3}
ggmap(melb) + geom_point(aes(x = Longitude, y = Latitude),
                         data = ped_loc, colour = "red", alpha = 0.7) +
              labs(title = "Melbourne Pedestrian Sensor Locations") +
              xlab(NULL) + ylab(NULL) +
              theme(plot.title = element_text(hjust = 0.5),
                    plot.caption = element_text(hjust = 0.5),
                    axis.text.x = element_blank(),
                    axis.text.y = element_blank(),
                    axis.ticks = element_blank())
```

@@

ggplot(original_df %>%  filter(Date_Time < "2014-07-06 00:00:00",
                               Date_Time > "2014-07-04 12:00:00")) +
  geom_line(aes(x = Date_Time, y = fit_dat_plot), colour = 'red',
            size = 2) +
  geom_line(aes(x = Date_Time, y = `Southern Cross Station`),
            size  = 2) +
  theme_minimal() + xlab("Time") +
  ylab("Pedestrian Counts") +
  ggtitle("Pedestrian Counts at Southern Cross Station on 04/07/2014 - 05/07/2014")
```

@@

- **Step 1:** Find which sensors have a "large" proportion of missing values  
- **Step 2:** At each sensor location, run appropriate GLM model with quasipoisson error distribution for imputation  
  * if "small", use GLM with `Hourly_Counts` as response variable and time and date based variables as predictors
  * if "large", use GLM with `Hourly_Counts` as response variable and neighbouring sensor counts as predictors
- **Step 3:** Replace missing values with imputed values  
---
## Step 1: Calculating proportion of missing values
- Need to first fix potential false zero-count values
  * Some observations of `Count = 0` may actually be missing values
  * Include a check for long periods with zero count, and replace with `NA` if too long
- Classify sensors as "small" or "large" proportion of missing values with a threshold proportion
  * We used 10% as the threshold proportion
```{r, fig.retina = 3, echo = FALSE, message = FALSE, fig.height = 4, fig.align = 'center'}
loc <- loc %>% mutate(cl = ifelse((loc$ids %in% small_miss_sensors),
                                  "Small", "Large"))
melbbw <- get_map(location=c(mean(range(ped_loc$Longitude)),
                            mean(range(ped_loc$Latitude))),
                            zoom=14, source = "stamen", maptype = "toner")

ggmap(melbbw) + geom_point(aes(x = lons, y = lats, colour = cl),
                         data = loc, alpha = .6, size = 4, shape = 15) +
              labs(title = "Melbourne Pedestrian Sensor Locations") +
              xlab(NULL) + ylab(NULL) +
              scale_color_viridis(name = "Proportion of missing values",
                                  discrete = TRUE) +
              theme(plot.title = element_text(hjust = 0.5),
                    plot.caption = element_text(hjust = 0.5),
                    axis.text.x = element_blank(),
                    axis.text.y = element_blank(),
                    axis.ticks = element_blank(),
                    legend.position = "bottom")
```

@@

>>>>

```{r, echo = FALSE, fig.height=5, fig.retina = 3}
ggmap(melb) + geom_point(aes(x = Longitude, y = Latitude),
                         data = ped_loc, colour = "red",
                         size = 4, alpha = 0.7, shape = 17) +
              labs(title = "Melbourne Pedestrian Sensor Locations") +
              xlab(NULL) + ylab(NULL) +
              theme(plot.title = element_text(hjust = 0.5),
                    plot.caption = element_text(hjust = 0.5),
                    axis.text.x = element_blank(),
                    axis.text.y = element_blank(),
                    axis.ticks = element_blank())
```

@@

ggplot(original_df %>%  filter(Date_Time < "2014-07-06 00:00:00",
                               Date_Time > "2014-07-04 12:00:00")) +
  geom_line(aes(x = Date_Time, y = fit_dat_plot),
            colour = plasma(1, begin = .75, end = 0),
            size = 2) +
  geom_line(aes(x = Date_Time, y = `Southern Cross Station`),
            size  = 2) +
  theme_minimal() + xlab("Time") +
  ylab("Pedestrian Counts") +
  ggtitle("Pedestrian Counts at Southern Cross Station on 04/07/2014 - 05/07/2014")
```

@@

- **Step 1:** Find which sensors have a "large" proportion of missing values  
- **Step 2:** At each sensor location, run appropriate GLM model with quasipoisson error distribution for imputation  
  * if "small", use GLM with `Hourly_Counts` as response variable and time and date based variables as predictors
  * if "large", use GLM with `Hourly_Counts` as response variable and neighbouring sensor counts as predictors
- **Step 3:** Replace missing values with imputed values  
---
## Step 1: Calculating proportion of missing values
- Need to first fix potential false zero-count values
  * Some observations of `Count = 0` may actually be missing values
  * Include a check for long periods with zero count, and replace with `NA` if too long
- Classify sensors as "small" or "large" proportion of missing values with a threshold proportion
  * We used 10% as the threshold proportion
```{r, fig.retina = 4, echo = FALSE, message = FALSE, fig.height = 4, fig.align = 'center', warning=FALSE}
loc <- loc %>% mutate(cl = ifelse((loc$ids %in% small_miss_sensors),
                                  "Small", "Large"))
melb2 <- get_map(location = c(mean(range(ped_loc$Longitude)),
                            mean(range(ped_loc$Latitude))),
                            zoom=14, maptype = "satellite")
perf_loc <- c(144.937, -37.826, 144.977, -37.794)

ggmap(melb2) + geom_point(aes(x = lons, y = lats, colour = cl),
                         data = loc, alpha = .6, size = 4, shape = 15) +
              labs(title = "Melbourne Pedestrian Sensor Locations") +
              xlab(NULL) + ylab(NULL) +
              scale_color_viridis(name = "Proportion of missing values",
                                  discrete = TRUE, option = "C") +
              scale_x_continuous(limits = c(144.937, 144.977)) +
              scale_y_continuous(limits = c(-37.83, -37.79)) +
              theme(plot.title = element_text(hjust = 0.5),
                    plot.caption = element_text(hjust = 0.5),
                    axis.text.x = element_blank(),
                    axis.text.y = element_blank(),
                    axis.ticks = element_blank(),
                    legend.position = "right")
```

@@

## Step 2: Replacing missing values with imputed values

- Need to fill the missing data at the sensors with small proportions of missing values first
  * The imputation models at sensors with large proportions of missing values are based on neighbouring sensors
  * Requires complete data at neighbours to work properly

```{r, echo = FALSE, fig.retina = 3, fig.width = 10, fig.height = 4}
fit_dat_plot <- data.frame(Date_Time = dfa2$Date_Time,
                           fit_ = dfa2$`Flagstaff Station`) %>%
                  filter(Date_Time < "2014-07-05 15:00:00",
                         Date_Time > "2014-07-03 23:00:00") %>%
                        .$fit_

p1 <-  ggplot(original_df %>%  filter(Date_Time < "2014-07-05 15:00:00",
                                 Date_Time > "2014-07-03 23:00:00")) +
        geom_line(aes(x = Date_Time, y = `Flagstaff Station`),
                  size  = 2) +
        theme_minimal() + xlab("Time") +
        ylab("Pedestrian Counts") +
        ggtitle("Flagstaff Station")
p2 <-  ggplot(original_df %>%  filter(Date_Time < "2014-07-05 15:00:00",
                                 Date_Time > "2014-07-03 23:00:00")) +
        geom_line(aes(x = Date_Time, y = fit_dat_plot),
                  colour = plasma(1, begin = .335, end = 0),
                  size = 2) +
        geom_line(aes(x = Date_Time, y = `Flagstaff Station`),
                  size  = 2) +
        theme_minimal() + xlab("Time") +
        ylab("Pedestrian Counts") +
        ggtitle("Flagstaff Station with imputed values")
grid.arrange(p1, p2, nrow = 1)
```
